---
title: "Electrical Workers Activity Recognition"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
author:
  - name: "Sahand Hajifar ^[Email: sahandha@buffalo.edu]"
    affiliation: Department of Industrial and Systems Engineering, University at Buffalo
    
bibliography: refs.bib
link-citations: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{chngpage}
  - \usepackage{caption}
  - \usepackage{chngpage}
  - \usepackage{color}
  - \usepackage[autostyle=true,english=american]{csquotes}
  - \usepackage{csvsimple}
  - \usepackage{framed}
  - \usepackage{graphicx}
  - \usepackage{hyperref}
  - \usepackage{lineno}
  - \usepackage{lscape}
  - \usepackage{mathptmx}
  - \usepackage{mathrsfs}
  - \usepackage{makecell}
  - \usepackage{mathtools}
  - \usepackage{media9}
  - \usepackage{multicol}
  - \usepackage{multirow}
  - \usepackage{secdot}
  - \usepackage{sectsty}
  - \usepackage{setspace}
  - \usepackage{subcaption}
  - \usepackage{tabulary}
  - \usepackage{titlesec}
  - \usepackage[colorinlistoftodos]{todonotes}
  - \usepackage{url}
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
    paged_df: TRUE
    code_download: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      out.width = "100%",
                      warning = FALSE,
                      message = FALSE) 
```

---

\newpage 
# Initialization

## R Initilization/Package Management 
The code snippet below presents the **R** packages used in our analysis. Additionally, the reader should note that we have used R version 3.6.3 in our analysis. The analysis is performed on an HP laptop equipped with a 2.10 GHz processor and 16.0 GB of RAM running a 64-bit version of Windows 10..
```{r error=FALSE, warning=FALSE, cache=FALSE, load_libraries, message=FALSE}
rm(list = ls()) # clear global environment
graphics.off() # close all graphics
#p_load is equivalent to combining both install.packages() and library()
pacman::p_load(tidyverse, magrittr, doParallel,
               caret, caretEnsemble, AUC, caTools, DMwR, MLmetrics,
               rpart, nnet, #rpart and multinom packages
               randomForest, e1071, foreach, import, #parRF packages
               kernlab, # svmRadial package
               conflicted,ggfortify, readxl, DMwR, ROSE, pander,
               wavelets, entropy) # Loading required packages and data

conflict_prefer("select", "dplyr") #Prefering dplyr::select over any other package
conflict_prefer("summarize", "dplyr") # similar to above but with dplyr::summarize

source("./00_customFunctions.R") # custom functions
set.seed(2020)
numCores = detectCores() -1 # to allow the use of other programs on machine
startTime = Sys.time()
```

## Data Extraction
In this section, acceleration data and annotation data are extracted from two separate folders ('Acc. Data' and 'Annotation Data').

```{r settings}
# subject_numbers = c(23,28,29,32,34,35,36,37,39,41)
subject_numbers = c(8,9,14,20,21,25,28,29,30,38)
# subject_numbers = 8:44
sub_range = 1:length(subject_numbers)
freq = 32 #Data collection frequency
```


```{r extract}
  Subject_Data <- sprintf("./Acc. Data/ACC%s.csv",subject_numbers) %>%
  lapply(read.csv) %>%
  lapply(function(x){x = x[-1,]; return(x)}) #remove first element which represents frequency

  #This extracts the start time of each subject based on unix time
  start_time = lapply(Subject_Data, function(x){temp = colnames(x)[1];return(temp)}) %>%
    lapply(function(x){temp = as.numeric(sub('.', '', x));return(temp)}) %>%
    unlist
  
  Subject_Data = lapply(Subject_Data, function(x) {colnames(x) = c('ax','ay','az'); x})
  
# Filtering
# n = 4
 fc = 10
 Fs = 32
# Wn = (2/Fs)*fc
# but_filt <- signal::butter(n=n, W=Wn, type="low")
# 
# Subject_Data = lapply(Subject_Data, function(x){
#   temp = apply(x, 2, signal::filter,filt = but_filt) ; temp})
  
subject_annotation = read_excel_allsheets("./Annotation Data/separate.xlsx")
  
  to_be_removed = !apply(matrix(names(subject_annotation)), 1, function(x){ ### removes the subjects that are not in 'subject_numbers'
     temp = x %in% (sprintf("P%s",subject_numbers)); temp
  })
  
subject_annotation[to_be_removed] = NULL
  
indices_list = vector(mode = "list", length = length(subject_annotation))
for (i in 1:length(indices_list)){
  indices_list[[i]] = freq * apply(matrix(subject_annotation[[i]][,2]), 1, time_diff, start_time = start_time[i])
  }
  
Names = apply(matrix(subject_annotation[[1]][,1]), 1, function(x){name = sub("_.*", "",x);return(name)})
```

## Information for Reproducing our Research

```{r reproduce}
    sInfo = sessionInfo()
    sInfo
```

# Visualization {.tabset .tabset-fade .tabset-pills}
In this section, acceleration plots are generated for different subjects. Red vertical lines represent the times that tasks start or end.

```{r acc_plot,echo=TRUE, results='asis'}
for (i in sub_range)
{
  Raw_Data<-ts(data = Subject_Data[[i]], start = 1, end = dim(Subject_Data[[i]])[1], frequency = 1, names =c("ax","ay","az") )
  cat(paste0("## Subject ",subject_numbers[i], " {-} \n"))
  
  Plot = ggplot2::autoplot(Raw_Data)+theme_bw()+ ggtitle(paste0("Subject ",toString(subject_numbers[i])))+ theme(plot.title = element_text(hjust = 0.5))
  
  Plot = Plot + geom_vline(xintercept = indices_list[[i]], colour="red")
  
  
  #for (j in 1:(length(indices_list))){
  # Plot = Plot + geom_vline(aes_(xintercept= indices_list[[i]][j],
  #               color=Names[j]))
  #}
  
  print(Plot)
  cat("\n \n")
}

```

# Feature Extraction
In this section, discrete wavelet transform (DWT) is applied to the acceleration data (aX, aY and aZ) in 4-second and 10-second time windows (numVars is the number of obs. per window). To get a better understanding of how the features are engineered, we refer the reader to our paper.

```{r feature_extract}
Names = Names[seq(2,length(Names),2)]
Names[Names == "ladder1" | Names == "ladder2"] = "ladder"
feature_list = vector(mode = "list", length = length(subject_annotation))

for (num_sub in 1:length(Subject_Data)){
  
  indices_list2 = vector(mode = "list", length = length(Names))
for (i in 1:length(indices_list2))
{
  tryCatch({
  indices_list2[[i]] = indices_list[[num_sub]][2*i-1]:indices_list[[num_sub]][2*i]
  }, error = function(e) {})
}
  
df = data.frame(acc= Subject_Data[[num_sub]]*9.81/64, label = NA)
for (i in 1:length(indices_list2)) { 
  rows = indices_list2[[i]]
  df[rows, 'label'] = Names[i]
}

df %<>% na.omit() # removing unlabeled acc observations
df$label %<>% as.factor() # converting the label to a factor

numVars = 128 # we can change that (number of datapoints per window)
numLevel = 5 # Number of levels from wavelet to extract

feature_X = featureX(df)
feature_Y = featureY(df)
feature_Z = featureZ(df)

feature = cbind(feature_X[,1:(dim(feature_X)[2]-1)],
                feature_Y[,2:(dim(feature_Y)[2]-1)],
                feature_Z[,2:(dim(feature_Z)[2])])

colnames(feature) = c('label', paste0("f", seq(1, 3*((numLevel+1)*12))), 'id')
feature_list[[num_sub]] = feature
  
}
save(feature_list, file = "feature_list.Rdata")
```
---

# References {-}